\chapter{Introduction}

The Climate Data Operators ({\CDO}) software is a collection of many operators
for standard processing of climate and forecast model output.
The operators include simple statistical and arithmetic functions, data selection
and subsampling tools, and spatial interpolation.
%{\CDO} is a collection of Operators to manipulate and analyse Climate model Data.
%There are a lot of operators for doing standard processing like simple statistic,
%arithmetic, selection and interpolation on datasets from climate and forecast models.
%The operators are primarily designed to aid manipulation and analysis of gridded scientific data
{\CDO} was developed to have the same set of processing functions for \htmlref{GRIB}{GRIB} and \htmlref{netCDF}{netCDF}
datasets in one package.

The Climate Data Interface (\htmlref{CDI}{CDI}) is used for the fast and file format
independent access to GRIB and netCDF datasets.
The local data formats SERVICE, EXTRA and IEG are also supported.

There are some limitations for GRIB and netCDF datasets.
A GRIB dataset must be consistent, similar to netCDF.
That means all time steps must have the same variables, and
within a time step each variable may occur only once.
NetCDF datasets are supported only with 2-dimensional, 3-dimensional and 4-dimensional
variables and the attributes should follow the
\href{http://ftp.unidata.ucar.edu/software/netcdf/docs/conventions.html}
     {GDT, COARDS or CF Conventions}.

The user interface and some operators are similar to the \htmlref{PINGO}{PINGO} package.
There are also some operators with the same name as in PINGO but with a 
different meaning. \htmlref{Appendix A}{appendixpingo} gives an overview of 
those operators.

The main {\CDO} features are:
\begin{itemize}
\item More than 400 operators available
\item Modular design and easily extendable with new operators
\item Very simple UNIX command line interface
\item A dataset can be processed by several operators,
      without storing the interim results in files % (No disk I/O).
\item Most operators handle datasets with missing values
\item Fast processing of large datasets
\item Support of many different grid types
\item Tested on many UNIX/Linux systems, Cygwin, and MacOS-X
%item Free available and runs on all UNIX platforms.
\end{itemize}

\input{install}

\input{usage}

\input{grid}

\input{zaxis}

\input{time}

\input{partab}

\input{missval}
