\chapter{Contributors}

\section{History}

{\CDO} was originally developed by Uwe Schulzweida at the Max Planck Institute for Meteorology (MPI-M).
The MPI-M, together with the DKRZ, has a long history in the development of tools for processing climate data.
{\CDO} was inspired by some of these tools, such as the PINGO package and the GRIB-Modules.

PINGO\footnote{Procedural INterface for GRIB formatted Objects} was developed by J\"urgen Waszkewitz,
 Peter Lenzen, and Nathan Gillet in 1995 at the DKRZ, Hamburg (Germany).
{\CDO} has a similar user interface and uses some of the PINGO routines.

The GRIB-Modules was developed by Heiko Borgert and Wolfgang Welke in 1991 at the MPI-M.
{\CDO} is using a similar module structure and also some of the routines.

\section{External sources}

{\CDO} has incorporated code from several sources:

\begin{description}
\item[afterburner] is a postprocessing program for ECHAM data and ECMWF analysis data,
originally developed by Edilbert Kirk, Michael Ponater and Arno Hellbach.
The afterburner code was modified for the {\CDO} opertors  {\htmlref{after}{after}},  {\htmlref{sp2gp}{sp2gp}},  {\htmlref{gp2sp}{gp2sp}}.

\item[SCRIP] is a software package used to generate interpolation weights for remapping
fields from one grid to another in spherical geometry \cite{SCRIP}.
It was developed at the Los Alamos National Laboratory by Philip W. Jones.
The SCRIP library was converted from Fortran to ANSI C and is used as
the base for the remapping operators in {\CDO}.

\item[YAC] (Yet Another Coupler) was jointly developed by DKRZ and MPI-M by Moritz Hanke and Rene Redler \cite{YAC}.
{\CDO} is using the clipping routines for the conservative remapping with {\htmlref{remapycon}{remapycon}}.

\item[libkdtree] a C99 implementation of the kd-tree algorithm developed by J\"org Dietrich.
\end{description}

{\CDO} uses tools from the GNU project, including automake, and libtool. 

\section{Contributors}

The primary contributors to the {\CDO} development have been: 

\begin{description}
\item[Uwe Schulweida]: Concept, design and implementation of {\CDO}, project coordination, and releases.

\item[Luis Kornblueh]: He supports {\CDO} from the beginning. 
His main contributions are GRIB performance and compression, GME and unstructured grid support.
Luis also helps with design and planning.

\item[Ralf M\"uller]: He is working on {\CDO} since 2009. 
His main contributions are the implementation of the User Portal, 
the ruby and python interface for all {\CDO} operators, the building process and the Windows support.
The {\CDO} User Portal was funded by the European Commission infracstructure project IS-ENES.
Ralf also helps a lot with the user support.
Implemented operators: {\htmlref{showunit}{showunit}}, {\htmlref{stdatm}{stdatm}}, {\htmlref{intlevel3d}{intlevel3d}}, {\htmlref{consecsum}{consecsum}},
{\htmlref{consects}{consects}}, {\htmlref{ngrids}{ngrids}}, {\htmlref{ngridpoints}{ngridpoints}}, {\htmlref{reducegrid}{reducegrid}}
% Webpage, ruby and python interface <------------------

\item[Cedrick Ansorge]: He worked on the {\CDO} software package as a student assistant at MPI-M from 2007-2011.
Implemented operators: {\htmlref{eof}{eof}}, {\htmlref{eof3d}{eof3d}}, {\htmlref{enscrps}{enscrps}}, {\htmlref{ensbrs}{ensbrs}},
 {\htmlref{maskregion}{maskregion}}, {\htmlref{bandpass}{bandpass}}, {\htmlref{lowpass}{lowpass}}, {\htmlref{highpass}{highpass}}, {\htmlref{smooth9}{smooth9}}

\item[Oliver Heidmann]: He worked on the {\CDO} software package as a student assistant at MPI-M from 2015-2018.

\item[Karin Meier-Fleischer]: She is working in the {\CDO} user support since 2017.

\item[Fabian Wachsmann]: He is working on {\CDO} for the CMIP6 project and is responsible for the operator cmor.

\item[Ralf Quast]: He worked on {\CDO} on behalf of the Service Gruppe Anpassung (SGA), DKRZ in 2006.
Implemented all ECA Indices of Daily Temperature and Precipitation Extremes, 
all percentile operators, module {\htmlref{YDRUNSTAT}{YDRUNSTAT}} and {\htmlref{wct}{wct}}.
% timcount, hourcount, daycount, moncount, yearcount, seascount

\item[Kameswarrao Modali]: Implemented operators: contour, shaded, grfill, vector, graph.

\item[Michal Koutek]: Implemented operators: {\htmlref{selmulti}{selmulti}}
{\htmlref{delmulti}{delmulti}}, {\htmlref{changemulti}{changemulti}}, {\htmlref{samplegrid}{samplegrid}},
{\htmlref{uvDestag}{uvDestag}}, {\htmlref{rotuvNorth}{rotuvNorth}}, {\htmlref{projuvLatLon}{projuvLatLon}}.

\item[Etienne Tourigny]: Implemented operators:  {\htmlref{setclonlatbox}{setclonlatbox}}, {\htmlref{setcindexbox}{setcindexbox}},
 {\htmlref{setvals}{setvals}}, {\htmlref{splitsel}{splitsel}}, {\htmlref{histfreq}{histfreq}}, {\htmlref{setrtoc}{setrtoc}}, {\htmlref{setrtoc2}{setrtoc2}}.

\item[Karl-Hermann Wieners]: Implemented operators: {\htmlref{aexpr}{aexpr}},  {\htmlref{aexprf}{aexprf}},  {\htmlref{selzaxisname}{selzaxisname}}.
\end{description}

% Frank Kaspar: Contributed to operator: import\_cmsaf

% Helmut Haak: Contributed to operator: adisit, adipot

% Martin Claus: Modified operator: vertwind (added missing value support)

% Ãlvaro M. Valdebenito: Modified operators:  settunits, settaxis and shifttime (added support for "seconds")

% Mathis Rosenhauer: Implemented operators: cmor (initial version)

%Reinhard Budich

\vspace{5mm}

Many users have contributed to {\CDO} by sending bug reports, patches
and suggestions over time.
Very helpful is also the active participation in the user forum of some users.
Here is an incomplete list:

\texttt{
\parbox{\textwidth}{
\mbox{Jaison-Thomas Ambadan},
\mbox{Harald Anlauf},
\mbox{Andy Aschwanden},
\mbox{Stefan Bauer},
\mbox{Simon Blessing},
\mbox{Renate Brokopf},
\mbox{Michael Boettinger},
\mbox{Tim Br\"ucher},
\mbox{Reinhard Budich},
\mbox{Martin Claus},
\mbox{Traute Cr\"uger},
\mbox{Irene Fischer-Bruns},
\mbox{Chris Fletscher},
\mbox{Helmut Frank},
\mbox{Kristina Fr\"ohlich},
\mbox{Oliver Fuhrer},
\mbox{Monika Esch},
\mbox{Pier Giuseppe Fogli},
\mbox{Beate Gayer},
\mbox{Veronika Gayler},
\mbox{Marco Giorgetta},
\mbox{David Gobbett},
\mbox{Holger Goettel},
\mbox{Helmut Haak},
\mbox{Stefan Hagemann},
\mbox{Angelika Heil},
\mbox{Barbara Hennemuth},
\mbox{Daniel Hernandez},
\mbox{Nathanael Huebbe},
\mbox{Thomas Jahns},
\mbox{Frank Kaspar},
\mbox{Daniel Klocke},
\mbox{Edi Kirk},
\mbox{Yvonne K\"ustermann},
\mbox{Stefanie Legutke},
\mbox{Leonidas Linardakis},
\mbox{Stephan Lorenz},
\mbox{Frank Lunkeit},
\mbox{Uwe Mikolajewicz},
\mbox{Laura Niederdrenk},
\mbox{Dirk Notz},
\mbox{Hans-J\"urgen Panitz},
\mbox{Ronny Petrik},
\mbox{Swantje Preuschmann},
\mbox{Florian Prill},
\mbox{Asela Rajapakse},
\mbox{Daniel Reinert},
\mbox{Hannes Reuter},
\mbox{Mathis Rosenhauer},
\mbox{Reiner Schnur},
\mbox{Martin Schultz},
\mbox{Dennis Shea},
\mbox{Kevin Sieck},
\mbox{Martin Stendel},
\mbox{Bjorn Stevens},
\mbox{Martina Stockhaus},
\mbox{Claas Teichmann},
\mbox{J\"org Trentmann},
\mbox{\'Alvaro M. Valdebenito},
\mbox{Geert Jan van Oldenborgh},
\mbox{Jin-Song von Storch},
\mbox{David Wang},
\mbox{Joerg Wegner},
\mbox{Heiner Widmann},
\mbox{Claudia Wunram},
\mbox{Klaus Wyser}
}
}

Please let me know if your name was omitted!
